---
class: 'article'
usepackage: '{fancyhdr}'
pagestye: '{fancy}'
fancyhead: '[C]{09/06/2023}'
fancyfoot: '[C]{\thepage}'
title: 'CSC8639 Interim Report: Explaining Time Series Downsampling'
author: 'Author: 210431461 | Supervisor: Matthew Forshaw'
citation_package: natbib
bibliography: 'bib.bib'
biblio-style: 'apalike'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
```

```{r ProjectTemplate, include = FALSE}
library(ProjectTemplate)
load.project()
```

The interim report is a short document which provides a brief overview of your proposed project and the progress you have made on it so far.

## Introduction (10%)

Decision-makers must trust that the data being considered sufficiently represents the situation they are deciding on. Trusting the data means trusting What data is selected, how the chosen data is collected and stored, and the capability of data practitioners to understand the quality, insights and limitations of this data. It is already difficult to determine and explain that the data reliably and truthfully reflects the situation in question across every phase of this data processing and analysis pipeline. The increasing volume of data today makes this more difficult [reference].

This problem is particularly pertinent to collections of observations obtained through repeated measurements over time @AusBureau, known as time series data. This data type is being "widely generated by industry and research at an increasing speed" @TVStore. This is putting unprecedented demand on resources (look up 6, 79 in @TVStore) and forcing data practitioners to select methods, such as aggregation, windowing, and downsampling, that reduce this volume to align with cost or time limitations, storage capabilities, and sustainability ambitions @Sveinn, @TVStore, @Shift. These reduction methods involve discarding data, which could result in the loss of important information and reduce the representativeness of the data.

However, reducing time series data is also a vital part of visualising this data for human observation. 

The research outlined by this interim report is exploring how to better explain the impact of downsampling time series data on its representativeness. It is hoped that this will help data practitioners confidently select their downsampling approach, better communicate the insights and limitations of downsampled data, and, in doing so, support decision-makers to trust the data they are considering. 

## Aim and Objectives (20%)


A clear statement of the high-level aim of the project and a list of scientific objectives. Remember that the project aim should be a short, high-level statement of the overall goal of the project or the hypothesis the project is investigating. The project objectives should be smaller key tasks or sub-goals that will support you in achieving your aim (i.e. if you accomplish all of your project objectives then you should have achieved the aim of the project). Remember that the project objectives should not be functional requirements for a system or low-level tasks.

## Overview of Progress (20%)

A brief description of the progress made so far (i.e. areas researched, analysis undertaken, software or model development undertaken).

## Project Plan (20%)

A timed work plan for the completion of the project. This should cover the complete duration of the project and provide clear scheduling information for the key project tasks. Note the plan must be visual and be based on a graphical representation of the project timeline (e.g. a Gantt chart). Tools such as teamgantt.comLinks to an external site. provide strong functionality and a free trial version.

## References (10%)

The report should contain an appropriate list of references that you have considered during your research and these should be cited appropriately within your report. These should be formatted appropriately and should be cited within the document.

## Data Management Plan (10%)

A data management plan (DMP) is a written document outlining how you are planning to manage your research data both during and after your research project. The plan should address what types of data will be collected and how the data will be documented, stored, shared and preserved.

## Form (10%)