---
class: 'article'
usepackage: '{fancyhdr}'
pagestye: '{fancy}'
fancyhead: '[C]{09/06/2023}'
fancyfoot: '[C]{\thepage}'
title: 'CSC8639 Interim Report: Explaining Time Series Downsampling'
author: 'Author: 210431461 | Supervisor: Matthew Forshaw'
citation_package: natbib
bibliography: 'bib.bib'
biblio-style: 'apalike'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
```

```{r ProjectTemplate, include = FALSE}
library(ProjectTemplate)
load.project()
```

The interim report is a short document which provides a brief overview of your proposed project and the progress you have made on it so far.

## Introduction (10%)


This should set the scene for the project and motivate why the project is interesting. Remember that this section should include citations to appropriate supporting references.

## Aim and Objectives (20%)

A clear statement of the high-level aim of the project and a list of scientific objectives. Remember that the project aim should be a short, high-level statement of the overall goal of the project or the hypothesis the project is investigating. The project objectives should be smaller key tasks or sub-goals that will support you in achieving your aim (i.e. if you accomplish all of your project objectives then you should have achieved the aim of the project). Remember that the project objectives should not be functional requirements for a system or low-level tasks.

## Overview of Progress (20%)

A brief description of the progress made so far (i.e. areas researched, analysis undertaken, software or model development undertaken).

## Project Plan (20%)

A timed work plan for the completion of the project. This should cover the complete duration of the project and provide clear scheduling information for the key project tasks. Note the plan must be visual and be based on a graphical representation of the project timeline (e.g. a Gantt chart). Tools such as teamgantt.comLinks to an external site. provide strong functionality and a free trial version.

## References (10%)

The report should contain an appropriate list of references that you have considered during your research and these should be cited appropriately within your report. These should be formatted appropriately and should be cited within the document.

## Data Management Plan (10%)

A data management plan (DMP) is a written document outlining how you are planning to manage your research data both during and after your research project. The plan should address what types of data will be collected and how the data will be documented, stored, shared and preserved.

## Form (10%)